Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations

### 2020-08-28 paper.Rnw
resolving merge conflict

The concept of LIME is relatively simple: use an interpretable model to approximate a complex model in a local region. However, a practical implementation of LIME is not straightforward, and research is being done to improve the procedure \citep{laugel:2018}. The current implementations of LIME \citep{ribeiro:2020} \kgc{Is there a better way to cite a Python package?} \hh{might be good to model the python package reference after the R package references, have a loog at reference limepy} \kgc{I kept thinking about this, and since the Python pakcgae doesn't have a manual, I ended up combining your limepy reference with the second second suggestion from https://academia.stackexchange.com/questions/14010/how-do-you-cite-a-github-repository. For now, I listed Ribeiro as the author, because he is the main contributor, but there are 40 other contributors to the package. Is it okay to just list Ribeiro?} \citep{pedersen:2020} offer various tuning parameters (see Section \ref{background}) that affect the explainer model and ultimately, the explanation. Since the explainer model is an approximation of the complex model and not a direct interpretation, the explanations produced by an explainer model are subject to the quality of the approximation. Thus, in order to achieve accurate explanations, the tuning parameters selected need to be assessed.

@@

All runs of LIME in this paper are executed in a forked version\footnote{https://github.com/goodekat/lime} of the R package \emph{lime} (version \Sexpr{packageVersion("lime")}) by \kge{\citet{pedersen:2020}}. The forked version is functionally indistinguishable from \kge{\citeauthor{pedersen:2020}'s} implementation but allows us to export internal values relevant for an assessment of the explainer. The diagnostic plots included in this paper are created using the R package \emph{limeaid} \citep{goode:2020}.

@@

We apply LIME using six sets of tuning parameters to all observations in the \data \ test set to observe variability across tuning parameters. A quantile bin based simulation method \kge{(samples are simulated uniformly from a specified number of quantile bins)} is used for five of the LIME applications with the number of bins varying from 2 to 6 by application. We use 6 bins as the maximum, because the complexity of the explanations increases with the number of bins. \kge{Note that 4 quantile bins are the default method in \emph{lime}.} The sixth application of LIME uses a kernel density simulation method \kge{(samples are drawn from kernel density approximations of the feature distributions)}. The default methods for feature selection (forward selection) and the computation of the weights (Gower distance raised to an exponent of 1) are used for all applications. (See Appendix~\ref{lime-details} for descriptions of the tuning parameters.)

@@

plot_grid(sine_poi_exp, exp_scatter) 
@
\caption{(left) Visualization from the \emph{lime} R package of \kge{a} \sout{the} LIME explanation for the \data \ prediction of interest \kge{identified in \autoref{fig:sine-plot-data}}. \sout{The bars represent the explainer model coefficients.} (right) \emph{Explanation Scatterplot}. The points represent the simulated data colored by the random forest probabilities and sized by the assigned LIME weights. \hh{XXX could you make the complex prediction **much** lighter for this example? The diamond and the lines are hard to see, but should be the focus.} \kgc{Is that light enough? Feel free to edit the alpha value in the plot\_explain\_scatter function as you feel necessary. I also made the lines depicting the explainer model wider in limeaid, and there is an option 'line\_size' that can be adjusted.} The prediction of interest is shown as the diamond shaped point. The explainer model indicator variables associated with $x_1$ and $x_2$ (quantile bins containing the prediction of interest) are depicted by the solid lines. The line colors represent the explainer model coefficient signs and indicate whether the feature supports a prediction of `blue' (blue lines) or supports a prediction of `red' (red lines). This figure shows that the quantile bins are not flexible enough to capture the relationship between the random forest predictions and the $x_1$ and $x_2$ values.}
\label{fig:sine-exp-scatter}
\end{figure*}

@@

\kgc{I really like the idea of separating the description of the \emph{lime} figure from the explanation scatterplot, but the more I think about it, I don't like that it get it's own sub-section. I think it would be nicer to allocate the sub-sections to the visual diagnostic plots. Instead, I switched to using a heading like you created for introducing the sine data. I also changed the wording of the heading a bit. What do you think? Another option would be to drop the heading altogether and flow into the next paragraph.} \hh{XXX I've just pulled the description of the left hand side of figure 4 out of the next section.}

\kge{Before introducing the visual diagnostics, let us consider a commonly used visualization of a LIME explanation that is shown in Figure \ref{fig:sine-exp-scatter} (left). The plot is created using the \emph{lime} R package and provides a visual representation of a LIME explanation for the prediction of interest indicated in \autoref{fig:sine-plot-data}. In particular, the explanation is obtained using 4 quantile bins to simulate the data.} \kgc{I moved the next few sentences here from the sub-section.} In this scenario, LIME converts continuous predictor variables to indicator variables identifying whether the variable value falls in the same quantile bin as the prediction of interest or not. The indicator variables are used as the explainer model features. \sout{To demonstrate a LIME explanation, we focus on the misclassified point indicated in Figure \ref{fig:sine-plot-data} by the diamond as the prediction of interest. The left plot in Figure \ref{fig:sine-exp-scatter} shows a visualization of the explanation for the prediction of interest as suggested by Ribeiro et al and implemented in the \emph{lime} R package.}

The \kge{plot} \sout{figure} \kge{reports \Sexpr{round(sine_poi_lime_default$label_prob[1], 2)} as} \sout{shows that} the random forest \sout{returned a} probability \sout{of \Sexpr{round(sine_poi_lime_default$label_prob[1], 2)}} that the observation of interest belongs to category blue {(denoted as the ``label" in the \kge{plot \sout{figure}}). The lengths of the bars represent the \kge{\sout{explainer model}} coefficients \kge{from a ridge regression model (the explainer model) fit using the R package \emph{glmnet} \citep{simon:2011}} associated with the indicator variables chosen \kge{via feature selection \sout{by LIME as important}}. The color of \kge{a \sout{the}} bar denotes the sign of the coefficient and whether the feature "supports" or "contradicts" a random forest classification of `blue'. The "explanation fit" \kge{\sout{value}} is the deviance ratio from \kge{\sout{the R package}} \emph{glmnet} \kge{\sout{\citep{simon:2011} for a ridge regression model}}. In other words, this is the $R^2$ value associated with the explainer model. In this case, it is \Sexpr{round(sine_poi_lime_default$model_r2[1], 2)} suggesting that the explainer model is not a good linear fit. However, it is commonly accepted that $R^2$ has limitations for assessing the quality of fit of a model \citep{sapra:2014}\kge{\sout{,}} and \kge{\sout{it}} should not be used as the only metric in a model assessment.

The explanation for the prediction of interest depicted in Figure \ref{fig:sine-exp-scatter} \kge{is} \sout{can be} interpreted as follows. A random forest classification of blue is supported by the prediction of interest having a value of $x_2$ that is greater than \Sexpr{str_split(sine_poi_lime_default$feature_desc[[1]], pattern = " ")[[1]][1]}, but the prediction of interest having a value of $x_1$ that is greater than \Sexpr{str_split(sine_poi_lime_default$feature_desc[[2]], pattern = " ")[[1]][1]} and less than or equal to \Sexpr{str_split(sine_poi_lime_default$feature_desc[[2]], pattern = " ")[[1]][5]} provides support against a classification of blue. \kge{Since the weight associated with $x_2$ has a larger magnitude than the weight associated with $x_1$, LIME explains that $x_2$ plays a more important role in the random forest prediction.} \kge{Additionally, consider that since the support for blue by $x_2$ outweights the support for red by $x_1$, the explainer model overall favors a label of blue for the prediction of interest, which} agrees with the random forest probability of \Sexpr{sine_poi_lime_default$label_prob[1]} for blue. Note that both models classify the observation incorrectly.

@@

\subsection{Explanation Scatterplots} \label{exp-scatter}
\hh{XXX I think we could reduce some of the example for the specific example a bit, now that the LIME explanation is moved out, there is not that much need to repeat the general ideas.} \kgc{I commented out the text that I think you were referring to. See what you think.}

\kgc{I moved the next sentence here from the previous paragraph.} Based on the plot \hh{on the left of \autoref{fig:sine-exp-scatter}} alone, it is not possible to make an informed assessment of the explanation. \hh{For a further assessment of the explainer model}, we turn to an \emph{explanation scatterplot}: \sout{the \emph{explanation scatterplot} is} a visual diagnostic for assessing the LIME claims of locality and fidelity for an individual explanation by juxtaposing the complex and explainer models in one plot. The format of an explanation scatterplot depends on the LIME simulation method. We introduce the explanation scatterplot here under the \emph{lime} R package default method \kge{four quantile bins.} \sout{in which the data are simulated uniformly from four quantile bins.}

The explanation scatterplot is built by plotting the LIME simulated data for the top two features identified by the explanation in a scatterplot and coloring these points by the  predictions from the complex model. The \kge{point} size \sout{of the points} represents the weight assigned by LIME.
\hh{In order to show the LIME results for the observation of interest,} lines \hh{are drawn on top of the points. These lines } represent the boundaries of the indicator variables used to fit the explainer model. The \kge{line} color \sout{of the lines} denote\kge{s} whether LIME indicates that a feature supports or contradicts a class prediction. Appendix \ref{exp-scatter_plus} addresses the explanation scatterplot formats in other LIME simulation scenarios.

\hh{An explanation scatterplot} \sout{for the} \kge{corresponding to the LIME explanation depicted on the left in \autoref{fig:sine-exp-scatter}} \sout{example of the previous section} \hh{is shown on the right hand side of \autoref{fig:sine-exp-scatter}.} %The points are the simulated values of $x_2$ versus $x_1$ colored by the random forest predictions and sized by the proximity weight used to fit the ridge regression explainer model. The explainer model is represented by the solid horizontal and vertical lines identifying the boundaries of the quantiles that contain the prediction of interest. The color of the lines indicate whether the explainer model coefficient is positive (blue) or negative (red). Thus, the $x_2$ coefficient is positive, so the $x_2$ value of the prediction of interest supports a random forest prediction of `blue', and the $x_1$ coefficient is negative, so the $x_1$ value of the prediction of interest contradicts a random forest prediction of `blue'. The prediction of interest is represented by the diamond shaped point.
By juxtaposing the random forest predictions and the explainer model boundaries, we are able to assess the faithfulness and localness of the explainer model.

First, consider the claim of localness. The weights \kge{decay relatively slowly} \sout{remain relatively high} outside of the intersection of the two quantile bins suggesting that the LIME explanation is highly influenced by points outside of the bin\kge{s} containing the prediction of interest. \sout{While i} \kge{However,} it is \sout{still} difficult to say \kge{if} \sout{whether or not} the claim of localness has been violated\kge{, because a definition of a local region is not specified. Depending a the region a viewer considers to be local, an argument could be made in favor of or against a violation of localness. The explanation scatterplot raises awareness of the unclear definition of a local region with LIME. Nevertheless,} it is possible to say that the weights assigned to the simulated data do not \kge{capture} \sout{agree with} the local region \kge{identified} \sout{assigned} by the intersection of the quantile bins.

\kgc{I tried rewriting this paragraph to make it less "technical". See what you think.} Now, consider the claim of faithfulness. \kge{It can be said that the majority of the points in the $x_2$ quantile support a prediction of blue, which is captured by the bar supporting a prediction of blue in LIME explanation, and a similar statement can be said about the $x_1$ quantile bin. These statements validate the explanation produced by LIME. However, the explanation scatterplot plot shows that the random forest does a good job of capturing the sine curve classification boundary by creating various sized rectangles with similar probabilities, which the LIME explanation does not pick up on. Thus, LIME does provide an explanation for the prediction, but it is a very poor explanation in terms of faithfulness to the random forest prediction regions.}

\sout{Within the intersection of the $x_1$ and $x_2$ quantile bins, there are more random forest probabilities above 0.5 \kge{than below}. This indicates why the magnitude of the coefficient associated with $x_2$ is larger than that of $x_1$. However, the pattern of the random forest predictions in this plot suggest that better explanations for this prediction exist. One example of a better explanation is to say that the prediction of interest received a random forest prediction greater than 0.5 since the observation of interest falls in the region where $x_1$ is less than 2.5 and $x_2$ is greater 4.5. A more localized explanation is to say that the region around the prediction of interest shows that observations with $-0.3 < x_1 <= 2.5$ and $4.8 < x_2 <= 8.75$ are all assigned probabilities greater than 0.5. The procedure implemented by LIME is not flexible enough to capture either of these explanations.}

\sout{The explanation scatterplot shows issues with a definition of locality and an oversimplification of the random forest \kge{\sout{model}}. LIME does provide an explanation for the prediction of interest, which makes sense based on the quantile bins used. However, the explanation scatterplot identifies that better explanations for the prediction of interest exist.}

It is difficult to assess linearity from \kge{this} \sout{an} explanation scatterplot, but a residual plot of the explainer model could be used to check the linearity claim. See Appendix \ref{residual-plot} for the residual plot associated with the explanation considered in Figure \ref{fig:sine-exp-scatter}, which shows a violation of the linearity claim.

\kge{This example} \sout{The \data \ } explanation only includes two features. In situations where more than two features are included in a LIME explanation, the explanation scatterplots can be extended to a generalized pairs plot \citep{emerson:2013} that includes all pairwise combinations of features. Generalized pairs plots (and scatterplot matrices in general) have diminishing value as the number of features increase \citep{jensen:2011} \citep{sweller:2011}. Machine learning models are commonly fit using a large number of features, and therefore, a generalized pairs plot of explanation scatterplots for all features would be ineffective. However, when applying LIME, the user selects the number of features to return in the explanation. In the \emph{lime} R package, \citet{pedersen:2020} encourage users to select less than 10 features \sout{to include in the explanation}. \kge{As long as a small number of features are returned in the LIME explanation,} \sout{As a result, specifying a small number of features to include in a LIME explanation makes} it \kge{is} feasible to use \kge{a} generalized pairs plot of explanation scatterplots. \kge{An example is shown in Section \ref{bullet-assess-ex}.}

@@

@
\caption{Hypothetical examples of feature heatmaps in two possible situations. The heatmaps show the top feature chosen for 10 cases across 5 different sets of \sout{LIME} tuning parameters. The color of the cell indicates the feature chosen by LIME. \kge{Situation 1 is the ideal, because the explanations are not dependent on tuning parameter values.}}
\label{fig:concept-feat-heat}

@@

<<sine-feat-heat, out.width = '0.45\\textwidth', fig.width = 4, fig.height = 4>>=
plot_feature_heatmap(sine_lime_explain$explain,
                order_method = "PCA") +
  theme_bw(base_family = "Times", base_size = 10) +
  theme(legend.position = "bottom",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.background = element_rect(color = "white", fill = "white")) +
  labs(y = "Case")
@
\caption{\emph{Feature Heatmap}. An example feature heatmap of the explanations from the applications of LIME with different tuning parameters to the \data \ test set. The cases from the test set are plotted on the y-axis, and the tuning parameter values (simulation methods here) are included on the x-axis. The colors of the tiles indicate the feature selected by LIME for the corresponding case and tuning parameter value. The plot is faceted by the first \kge{and second most important} features selected by LIME (top \kge{and bottom} facet\kge{s, respectively}) \sout{ and the second most important features (bottom facet).} \kge{The vertical facets separate the} \sout{The column facets separate the} kernel density method from the quantile bin methods. The vertical striping indicates that the LIME explanations are not consistent across tuning parameters.}
\label{fig:sine-feat-heat}
\end{figure}

Explanations produced by LIME are likely to be affected by the choice of tuning parameters. \sout{An} \kge{A hypothetical} example of this is shown by Figures \ref{fig:concept-plot-good} and \ref{fig:concept-plot-bad} where the method used to weight the observations influenced the explanation. As of the time of writing this manuscript, we have encountered no recommendations for how to \kge{specify the parameter values} \sout{determine which method to use} besides for the default settings in \emph{lime}. In order to compare the explanations produced by LIME using different tuning parameters \kge{and provide another perspective for assessing localness}, we visualize an overview \kge{of} the explanations \kge{with} \sout{in} the \emph{feature heatmap} diagnostic plot.

The feature heatmap uses colors to identify the features selected by LIME \kge{across multiple predictions (referred to as cases here) and tuning paramter values} \sout{across tuning parameters for sets of explanations (referred to as cases here)} \kge{organized} \sout{faceted} by the \sout{position of} feature importance assigned by LIME. That is, for LIME applied with $t$ sets of tuning parameters to $n$ cases to select the $f$ top features, create $f$ heatmaps (one for each of the positions of importance determined by the magnitude of the explainer model coefficients) with the cases on the y-axis, the tuning parameters on the x-axis, and the cells colored by the feature chosen for the corresponding case and tuning parameter. Additional tuning parameters may also be included in the plot via facets. \sout{This plot is used assess the locality of the explanations and to compare results from different input options.}

Two hypothetical examples of feature heatmaps are included in \autoref{fig:concept-feat-heat}. The plots \kge{are} \sout{were} created with the assumption that LIME is applied to select the top feature out of $p=\kge{4}$ \sout{8} features for $n=10$ cases with $t=5$ sets of tuning parameters. Situation 1 \kge{is} \sout{shows} an example where the features selected are consistent across tuning parameters within a case but vary across cases within a tuning parameter. This is the ideal situation, because the LIME explanations do not depend on the tuning parameters but do depend on the location of the observation in the feature space. Situation 2 \kge{is} \sout{shows} an example where the features selected vary across tuning parameters within a case but are consistent across cases within a tuning parameter. This situation indicates that the features selected by LIME are dependent on the tuning parameters, and the explanations \kge{may not be} \sout{are not} local, because the same feature is chosen regardless of the case. In practice, it is expected that the plot will exhibit a combination of these two situations.

\autoref{fig:sine-feat-heat} shows \kge{a} \sout{the} feature heatmap for the LIME applications to the 100 observations in the \data \ test set. The most important \kge{and second most important features selected by lime are showin in the top and bottom facets, respectively.} \sout{(top facet) and second most important (bottom facet) features are shown in this plot.} For the quantile bins, the original features prior to the indicator variable transformation are included since it is obvious that different features would be selected when the sizes of the bins change. This figure shows that for kernel density and 2 quantile bins, LIME selects $x_1$ as the most important feature and $x_2$ as the second most important feature across all cases in the test set. These explanations are not local. There is variability in the features selected by LIME for 3 to 6 quantile bins \kge{suggesting more local explanations}. \sout{For these tuning parameters, there is a mix of horizontal and vertical stripes,} \kge{There are signs of vertical striping,} which suggests a dependence on tuning parameters. Note that the explanations from 3 and 5 quantile bins include the selection of the random noise variable ($x_3$) as an important variable in many predictions, which should not be the case. The pattern seen in the explanations for 3 to 6 quantile bins may suggest local explanations, but the dependence on tuning parameter makes it unclear which set of explanations to use.

@@

@
\caption{\emph{Assessment metric plot}. The assessment metrics computed on the applications of LIME to the \data \ test set are included in this example assessment metric plot. Each horizontal facet corresponds to one of three metrics: average $R^2$, average fidelity, or MSEE. The LIME tuning parameter \kge{values} (simulation methods in this example) \kge{are} \sout{is} plotted on the x-axis, and the metric values are shown on the y-axis. The points are colored by rank (within a metric) indicating best to worst (dark to light). The kernel density simulation methods performs well \kge{across} \sout{according to} all three metrics.}
\label{fig:sine-assess-metric}
\end{figure}

@@

\hh{A taxonomy of author contributions is suggested by Elsevier - I put the link in an email}

>>>>

The concept of LIME is relatively simple: use an interpretable model to approximate a complex model in a local region. However, a practical implementation of LIME is not straightforward, and research is being done to improve the procedure \citep{laugel:2018}. The current implementations of LIME \citep{ribeiro:2020} \kgc{Is there a better way to cite a Python package?} \hh{might be good to model the python package reference after the R package references, have a loog at reference limepy} \kgc{I kept thinking about this, and since the Python pakcgae doesn't have a manual, I ended up combining your limepy reference with the second second suggestion from https://academia.stackexchange.com/questions/14010/how-do-you-cite-a-github-repository. For now, I listed Ribeiro as the author, because he is the main contributor, but there are 40 other contributors to the package. Is it okay to just list Ribeiro?}\hh{I would go with 'and contributors' - if we just cite Ribeiro people might not realize that it's truly not just him.} \citep{pedersen:2020} offer various tuning parameters (see Section \ref{background}) that affect the explainer model and ultimately, the explanation. Since the explainer model is an approximation of the complex model and not a direct interpretation, the explanations produced by an explainer model are subject to the quality of the approximation. Thus, in order to achieve accurate explanations, the tuning parameters selected need to be assessed.

@@

All runs of LIME in this paper are executed in a forked version\footnote{https://github.com/goodekat/lime} of the R package \emph{lime} (version \Sexpr{packageVersion("lime")}) by \citet{pedersen:2020}. The forked version is functionally indistinguishable from \citeauthor{pedersen:2020}'s implementation but allows us to export internal values relevant for an assessment of the explainer. The diagnostic plots included in this paper are created using the R package \emph{limeaid} \citep{goode:2020}.

@@

We apply LIME using six sets of tuning parameters to all observations in the \data \ test set to observe variability across tuning parameters. A quantile bin based simulation method (samples are simulated uniformly from a specified number of quantile bins) is used for five of the LIME applications with the number of bins varying from 2 to 6 by application. We use 6 bins as the maximum, because the complexity of the explanations increases with the number of bins. Note that 4 quantile bins are the default method in \emph{lime}. The sixth application of LIME uses a kernel density simulation method (samples are drawn from kernel density approximations of the feature distributions). The default methods for feature selection (forward selection) and the computation of the weights (Gower distance raised to an exponent of 1) are used for all applications. (See Appendix~\ref{lime-details} for descriptions of the tuning parameters.)

@@

plot_grid(sine_poi_exp, exp_scatter) 
@
\caption{(left) Visualization from the \emph{lime} R package of a LIME explanation for the \data \ prediction of interest identified in \autoref{fig:sine-plot-data}. (right) \emph{Explanation Scatterplot}. The points represent the simulated data colored by the random forest probabilities and sized by the assigned LIME weights. %\hh{XXX could you make the complex prediction **much** lighter for this example? The diamond and the lines are hard to see, but should be the focus.} \kgc{Is that light enough? Feel free to edit the alpha value in the plot\_explain\_scatter function as you feel necessary. I also made the lines depicting the explainer model wider in limeaid, and there is an option 'line\_size' that can be adjusted.} \hh{thank you!}
The prediction of interest is shown as the diamond shaped point. The explainer model indicator variables associated with $x_1$ and $x_2$ (quantile bins containing the prediction of interest) are depicted by the solid lines. The line colors represent the explainer model coefficient signs and indicate whether the feature supports a prediction of `blue' (blue lines) or supports a prediction of `red' (red lines). This figure shows that the quantile bins are not flexible enough to capture the relationship between the random forest predictions and the $x_1$ and $x_2$ values.}
\label{fig:sine-exp-scatter}
\end{figure*}

@@

%\kgc{I really like the idea of separating the description of the \emph{lime} figure from the explanation scatterplot, but the more I think about it, I don't like that it get it's own sub-section. I think it would be nicer to allocate the sub-sections to the visual diagnostic plots. Instead, I switched to using a heading like you created for introducing the sine data. I also changed the wording of the heading a bit. What do you think? Another option would be to drop the heading altogether and flow into the next paragraph.} \hh{XXX I've just pulled the description of the left hand side of figure 4 out of the next section.}
\hh{XXX I like the idea of not giving this section an official header, but I like the way it looks like now with the bold header - that should help the reader what's going on.}

Before introducing the visual diagnostics, let us consider a commonly used visualization of a LIME explanation that is shown in Figure \ref{fig:sine-exp-scatter} (left). The plot is created using the \emph{lime} R package and provides a visual representation of a LIME explanation for the prediction of interest indicated in \autoref{fig:sine-plot-data}. In particular, the explanation is obtained using 4 quantile bins to simulate the data. %\kgc{I moved the next few sentences here from the sub-section.} 
In this scenario, LIME converts continuous predictor variables to indicator variables identifying whether the variable value falls in the same quantile bin as the prediction of interest or not. The indicator variables are used as the explainer model features.

The plot  reports \Sexpr{round(sine_poi_lime_default$label_prob[1], 2)} as  the random forest  probability  that the observation of interest belongs to category blue (denoted as the ``label" in the plot). The lengths of the bars represent the  coefficients from a ridge regression model (the explainer model) fit using the R package \emph{glmnet} \citep{simon:2011} associated with the indicator variables chosen via feature selection. The color of a  bar denotes the sign of the coefficient and whether the feature "supports" or "contradicts" a random forest classification of `blue'. The "explanation fit" is the deviance ratio from  \emph{glmnet}. In other words, this is the $R^2$ value associated with the explainer model. In this case, it is \Sexpr{round(sine_poi_lime_default$model_r2[1], 2)} suggesting that the explainer model is not a good linear fit. However, it is commonly accepted that $R^2$ has limitations for assessing the quality of fit of a model \citep{sapra:2014} and  should not be used as the only metric in a model assessment.

The explanation for the prediction of interest depicted in Figure \ref{fig:sine-exp-scatter} is interpreted as follows. A random forest classification of blue is supported by the prediction of interest having a value of $x_2$ that is greater than \Sexpr{str_split(sine_poi_lime_default$feature_desc[[1]], pattern = " ")[[1]][1]}, but the prediction of interest having a value of $x_1$ that is greater than \Sexpr{str_split(sine_poi_lime_default$feature_desc[[2]], pattern = " ")[[1]][1]} and less than or equal to \Sexpr{str_split(sine_poi_lime_default$feature_desc[[2]], pattern = " ")[[1]][5]} provides support against a classification of blue. Since the weight associated with $x_2$ has a larger magnitude than the weight associated with $x_1$, LIME explains that $x_2$ plays a more important role in the random forest prediction. Additionally, consider that since the support for blue by $x_2$ outweights the support for red by $x_1$, the explainer model overall favors a label of blue for the prediction of interest, which agrees with the random forest probability of \Sexpr{sine_poi_lime_default$label_prob[1]} for blue. Note that both models classify the observation incorrectly.

@@

\subsection{Explanation Scatterplots} \label{exp-scatter}
%\hh{XXX I think we could reduce some of the example for the specific example a bit, now that the LIME explanation is moved out, there is not that much need to repeat the general ideas.} \kgc{I commented out the text that I think you were referring to. See what you think.}

%\kgc{I moved the next sentence here from the previous paragraph.} 
Based on the plot on the left of \autoref{fig:sine-exp-scatter} alone, it is not possible to make an informed assessment of the explanation. For a further assessment of the explainer model, we turn to an \emph{explanation scatterplot}:  a visual diagnostic for assessing the LIME claims of locality and fidelity for an individual explanation by juxtaposing the complex and explainer models in one plot. The format of an explanation scatterplot depends on the LIME simulation method. We introduce the explanation scatterplot here under the \emph{lime} R package default method of four quantile bins.

The explanation scatterplot is built by plotting the LIME simulated data for the top two features identified by the explanation in a scatterplot and coloring these points by the  predictions from the complex model. The point size  represents the weight assigned by LIME.
In order to show the LIME results for the observation of interest, lines are drawn on top of the points. These lines  represent the boundaries of the indicator variables used to fit the explainer model. The line color  denotes whether LIME indicates that a feature supports or contradicts a class prediction. Appendix \ref{exp-scatter_plus} addresses the explanation scatterplot formats in other LIME simulation scenarios.

An explanation scatterplot corresponding to the LIME explanation depicted on the left in \autoref{fig:sine-exp-scatter} is shown on the right hand side of \autoref{fig:sine-exp-scatter}.
%The points are the simulated values of $x_2$ versus $x_1$ colored by the random forest predictions and sized by the proximity weight used to fit the ridge regression explainer model. The explainer model is represented by the solid horizontal and vertical lines identifying the boundaries of the quantiles that contain the prediction of interest. The color of the lines indicate whether the explainer model coefficient is positive (blue) or negative (red). Thus, the $x_2$ coefficient is positive, so the $x_2$ value of the prediction of interest supports a random forest prediction of `blue', and the $x_1$ coefficient is negative, so the $x_1$ value of the prediction of interest contradicts a random forest prediction of `blue'. The prediction of interest is represented by the diamond shaped point.
By juxtaposing the random forest predictions and the explainer model boundaries, we are able to assess the faithfulness and localness of the explainer model.

First, consider the claim of localness. The weights decay relatively slowly  outside of the intersection of the two quantile bins suggesting that the LIME explanation is highly influenced by points outside of the bins containing the prediction of interest. However, it is  difficult to say if the claim of localness has been violated, because a definition of a local region is not specified. Depending on what region a viewer considers to be local, an argument could be made in favor of or against a violation of localness. The explanation scatterplot raises awareness of the unclear definition of a local region with LIME. Nevertheless, it is possible to say that the weights assigned to the simulated data do not capture  the local region identified by the intersection of the quantile bins.

%\kgc{I tried rewriting this paragraph to make it less "technical". See what you think.}\hh{I like it - the arguments are also more to the point }
Now, consider the claim of faithfulness: It can be said that the majority of the points in the $x_2$ quantile support a prediction of blue, which is captured by the bar supporting a prediction of blue in the LIME explanation, and a similar statement can be made about the $x_1$ quantile bin. These statements validate the explanation produced by LIME. However, the explanation scatterplot plot shows that the random forest performs well at capturing the sine curve classification boundary by creating various sized rectangles consisting of predictions with similar probabilities, which the LIME explanation does not pick up on. Thus, LIME does provide an explanation for the prediction, but it is a very poor explanation in terms of faithfulness to the random forest prediction regions.

%\sout{Within the intersection of the $x_1$ and $x_2$ quantile bins, there are more random forest probabilities above 0.5 \kge{than below}. This indicates why the magnitude of the coefficient associated with $x_2$ is larger than that of $x_1$. However, the pattern of the random forest predictions in this plot suggest that better explanations for this prediction exist. One example of a better explanation is to say that the prediction of interest received a random forest prediction greater than 0.5 since the observation of interest falls in the region where $x_1$ is less than 2.5 and $x_2$ is greater 4.5. A more localized explanation is to say that the region around the prediction of interest shows that observations with $-0.3 < x_1 <= 2.5$ and $4.8 < x_2 <= 8.75$ are all assigned probabilities greater than 0.5. The procedure implemented by LIME is not flexible enough to capture either of these explanations.}

%\sout{The explanation scatterplot shows issues with a definition of locality and an oversimplification of the random forest \kge{\sout{model}}. LIME does provide an explanation for the prediction of interest, which makes sense based on the quantile bins used. However, the explanation scatterplot identifies that better explanations for the prediction of interest exist.}

It is difficult to assess linearity from this explanation scatterplot, but a residual plot of the explainer model could be used to check the linearity claim. See Appendix~\ref{residual-plot} for the residual plot associated with the explanation considered in Figure \ref{fig:sine-exp-scatter}, which shows a violation of the linearity claim.

This example explanation only includes two features. In situations where more than two features are included in a LIME explanation, the explanation scatterplots can be extended to a generalized pairs plot \citep{emerson:2013} that includes all pairwise combinations of features. Generalized pairs plots (and scatterplot matrices in general) have diminishing value when the number of features increase \citep{jensen:2011} \citep{sweller:2011}. Machine learning models are commonly fit using a large number of features, and therefore, a generalized pairs plot of explanation scatterplots for all features would be ineffective. However, when applying LIME, the user selects the number of features to return in the explanation. In the \emph{lime} R package, \citet{pedersen:2020} encourage users to select less than 10 features. As long as a small number of features are returned in the LIME explanation,  it is feasible to use a generalized pairs plot of explanation scatterplots. An example is shown in Section~\ref{bullet-assess-ex}.

@@

@
\caption{Hypothetical examples of feature heatmaps in two possible situations. The heatmaps show the top feature chosen for 10 cases across 5 different sets of  tuning parameters. The color of the cell indicates the feature chosen by LIME. Situation 1 is the ideal, because the explanations vary across cases but do not dependent on specific tuning parameter values.}
\label{fig:concept-feat-heat}

@@

<<sine-feat-heat, out.width = '0.45\\textwidth', fig.width = 4, fig.height = 4>>=
plot_feature_heatmap(sine_lime_explain$explain,
                order_method = "PCA") +
  theme_bw(base_family = "Times", base_size = 10) +
  theme(legend.position = "bottom",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.background = element_rect(color = "white", fill = "white")) +
  labs(y = "Case")
@
\caption{\emph{Feature Heatmap}. An example feature heatmap of the explanations from the applications of LIME with different tuning parameters to the \data \ test set. The cases from the test set are plotted on the y-axis, and the tuning parameter values (simulation methods here) are included on the x-axis. The colors of the tiles indicate the feature selected by LIME for the corresponding case and tuning parameter value. The plot is faceted by the first and second most important features selected by LIME (top and bottom facets, respectively). The vertical facets separate the  kernel density method from the quantile bin methods. The vertical striping indicates that the LIME explanations are not consistent across tuning parameters.}
\label{fig:sine-feat-heat}
\end{figure}

Explanations produced by LIME are likely to be affected by the choice of tuning parameters. A hypothetical example of this is shown by Figures \ref{fig:concept-plot-good} and \ref{fig:concept-plot-bad} where the method used to weight the observations influenced the explanation. As of the time of writing this manuscript, we have encountered no recommendations for how to specify the parameter values besides for the default settings in \emph{lime}. In order to compare the explanations produced by LIME using different tuning parameters and provide another perspective for assessing localness, we visualize an overview of the explanations with the \emph{feature heatmap} diagnostic plot.

The feature heatmap uses colors to identify the features selected by LIME across multiple predictions (referred to as cases here) and tuning paramter values organized by the  feature importance assigned by LIME. That is, for LIME applied with $t$ sets of tuning parameters to $n$ cases to select the $f$ top features, create $f$ heatmaps (one for each of the positions of importance determined by the magnitude of the explainer model coefficients) with the cases on the $y$-axis, the tuning parameters on the $x$-axis, and the cells colored by the feature chosen for the corresponding case and tuning parameter. Additional tuning parameters may also be included in the plot via facets.

Two hypothetical examples of feature heatmaps are included in \autoref{fig:concept-feat-heat}. The plots are created with the assumption that LIME is applied to select the top feature out of $p=\kge{4}$  features for $n=10$ cases with $t=5$ sets of tuning parameters. Situation 1 is  an example where the features selected are consistent across tuning parameters within a case but vary across cases within a tuning parameter. This is the ideal situation, because the LIME explanations do not depend on the tuning parameters but do depend on the location of the observation in the feature space. Situation 2 is an example where the selected features  vary across tuning parameters within a case but are consistent across cases within a tuning parameter. This situation indicates that the features selected by LIME are dependent on the tuning parameters, and the explanations may not be  local, because the same feature is chosen regardless of the case. In practice, it is expected that the plot will exhibit a combination of these two situations.

\autoref{fig:sine-feat-heat} shows a feature heatmap for the LIME applications to the 100 observations in the \data \ test set. The most important and second most important features selected by lime are showin in the top and bottom facets, respectively. For the quantile bins, the original features prior to the indicator variable transformation are included since it is obvious that different features would be selected when the sizes of the bins change. This figure shows that for kernel density and 2 quantile bins, LIME selects $x_1$ as the most important feature and $x_2$ as the second most important feature across all cases in the test set. These explanations are not local. There is variability in the features selected by LIME for 3 to 6 quantile bins suggesting more local explanations. There are signs of vertical striping, which suggests a dependence on tuning parameters. Note that the explanations from 3 and 5 quantile bins include the selection of the random noise variable ($x_3$) as an important variable in many predictions, which should not be the case. The pattern seen in the explanations for 3 to 6 quantile bins may suggest local explanations, but the dependence on tuning parameter makes it unclear which set of explanations to use.

@@

@
\caption{\emph{Assessment metric plot}. The assessment metrics computed on the applications of LIME to the \data \ test set are included in this example assessment metric plot. Each horizontal facet corresponds to one of three metrics: average $R^2$, average fidelity, or MSEE. The LIME tuning parameter values (simulation methods in this example) are plotted on the x-axis, and the metric values are shown on the y-axis. The points are colored by rank (within a metric) indicating best to worst (dark to light). The kernel density simulation methods performs well across all three metrics.}
\label{fig:sine-assess-metric}
\end{figure}

@@
